Ethnography of Online Performance Artists

### 2017-03-18 proposals/Brown_Magic_Grant_2017/main.tex
fixes based on jeff's notes

\title{Ethnography of Online Performance Artists}
\author{Ali Alkhatib \& Jeffrey Nagy
--- DRAFT \DTMcurrenttime~\today}
\date{\today}
\urlstyle{same}
\hypersetup{
  colorlinks,
  citecolor=Black,
  urlcolor=Blue}

@@

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[width=0.5\textwidth]{figures/cgpgrey.png}
\caption{\label{fig:Magic}A popular YouTuber discussing election results and various methods of determining election winners.}
\end{wrapfigure}
Video sharing sites like YouTube and Twitch
have captured the attention of billions of people around the world
in the span of little more than a decade.
Content producers
--- known informally as ``YouTubers'' and ``streamers'' ---
have begun to make their livings on these sites.
Indeed, these performers have come to rely on
the technical infrastructure these platforms provide as well as
the access to audiences that these platforms provide as well.
In doing so, these sites have in a real sense democratized media and entertainment industries by
making the avenues to large audiences substantially more accessible.
And thus, in these transformations into publics, what started as
a recreational hobby
has become a bustling hub for professional artists to
engage with audiences,
practice their craft, and 
hone their skills~(\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}).

But recently, tensions between these online performers and the platforms on which they work have shaken people's
assumptions regarding the nature of the space they're using.
Performers on now--defunct Vine called (unsuccessfully) to be better compensated~(\cite{vineWantsMoney,vineInsiderMeeting});
YouTubers meanwhile struggle to defend their unconventional performance work~(\cite{h3h3Lawsuit})
and individually and collectively field accusations of misconduct of various forms~(\cite{youtubeDramaResponses}).
These events highlight the politics of an emerging form of expression and art,
and specifically the sometimes controversial nature of the experimentation that performers engage in.

One issue frustrating this potential research area is that
the platforms have changed so dramatically in as few as 5 to 10 years that
it can seem as though researchers are trying to wade through shifting sands.
Scholarship on content creators loses bearing rapidly when
economic factors motivate aspiring professional entertainers to join the community~(\cite{Hamilton:2014:STF:2611105.2557048});
the tension between these entertainers experimenting with content remixing and
the holders of the copyright of the source material
(for example, see~(\cite{Hilderbrand48}))
shifts dramatically when
the platform's copyright enforcement mechanisms advance
not just incrementally but substantially~(\cite{kim2012institutionalization}).
And the popularity of candidate field sites can vary chaotically
(see, for example, Vine reaching as many as 200 million active users and
abrupt shutdown less than a year later~(\cite{vineDecline})).

@@

To that end, we can liken some aspects of online performance art on YouTube and Twitch
to street performance and busking,
allowing us to relate ostensibly new phenomena to robust scholarship.
As is the case with street performance,
online performance isn't restricted to classically trained professionals;
instead, performers (both online and in the streets) practice more experimental 
drawing on their candid interaction with their audiences,
which can change from one performance to the next.

This connection has purchase, but it also has gaps, mostly owing to the unique nature of the internet~(\cite{miller2011understanding}).
For one thing, most video sharing websites keep videos
in perpetuity
--- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
YouTube comments linger for years after they're made, forming in some sense ``networked publics''~(\cite{boyd2007youth}).
And these say nothing of the different nature of algorithmic enforcement of rules and policies, and
the near--perfect ability of these platforms to redirect revenue and even people's focus in an instant.

@@

\section*{Research Plan}
We propose to carry out qualitative research with major online content producers
on streaming platforms such as Twitch and YouTube, and video sharing sites like YouTube.
Much of this work will consist of ethnographic fieldwork, semi--structured interviews, and
open--ended exploration of the field site as we find parallels and divergences in the domain
of online performance art from street and other emergent formats of performance and entertainment.
A timeline of these phases is outlined here:

@@

We'll recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
--- in other words, people who have gone ``full--time''.
We'll aim for approximately two--thirds of our participants to be principally active on YouTube and one--third on Twitch, as
the former is significantly larger (and the community seemingly more diverse), suggesting
a need to get more varied perspectives from participants.

@@

% Each stage of research will focus on a specific thematic question for two reasons:
% first, to incrementally inform our subsequent, potentially more contentious, research questions; and
% second, to provide us with focused topics which we can scrutinize with the aforementioned historical lens of street performance.

>>>>

\title{Ethnography of Online Performance Artists}
\author{Ali Alkhatib \& Jeffrey Nagy
}
\date{\today}
\urlstyle{same}
\hypersetup{
  colorlinks,
  citecolor=Black,
  urlcolor=Blue}

@@

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\includegraphics[width=0.5\textwidth]{figures/cgpgrey.png}
\caption{\label{fig:Magic}A popular YouTuber discussing election results and various methods of determining election winners.}
\end{wrapfigure}
Video sharing sites like YouTube and Twitch
have captured the attention of billions of people around the world
in the span of little more than a decade.
Content producers
--- known informally as ``YouTubers'' and ``streamers'' ---
have begun to make their livings on these sites.
Indeed, these performers have come to rely on
the technical infrastructure these platforms provide as well as
the access to broad audiences and the promise of exposure.
In doing so, these sites have in a real sense democratized media and entertainment industries by
making the avenues to large audiences substantially more accessible.
And thus, in these transformations into Deweyan--style publics~(\cite{dewey2012public,disalvo2009design}), what started as
a recreational hobby
has become a bustling hub for professional artists to
engage with audiences,
practice their craft, and 
hone their skills~(\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}).

But recently, tensions between these online performers and the platforms on which they work have shaken
assumptions performers held regarding the nature of the space they're using,
revealing breakdowns in the expectations of these systems~(\cite{doi:10.1177/1461444809342738}).
Performers on now--defunct Vine called (unsuccessfully) to be better compensated~(\cite{vineWantsMoney,vineInsiderMeeting});
YouTubers meanwhile struggle to defend their unconventional performance work~(\cite{h3h3Lawsuit})
and individually and collectively field accusations of misconduct of various forms~(\cite{youtubeDramaResponses}).
These events highlight the politics of an emerging form of expression and art,
and specifically the sometimes controversial nature of the experimentation that performers engage in.

% One issue frustrating this potential research area is that
One issue research in this area must overcome is that
platforms have changed so dramatically in as few as 5 to 10 years that
it can seem as though researchers are trying to navigate through what seem at times like shifting sands.
Scholarship on content creators loses bearing rapidly when
economic factors motivate performers to turn a recreation or a hobby
--- such as video game streaming~(\cite{Hamilton:2014:STF:2611105.2557048}) --- into a career;
the tension between these entertainers experimenting with content remixing and
the holders of the copyright of the source material
(for example, see~(\cite{Hilderbrand48}))
shifts dramatically when
the platform's copyright enforcement mechanisms advance
not just incrementally but substantially~(\cite{kim2012institutionalization}).
And the popularity of candidate field sites can vary chaotically
(see, for example, Vine reaching as many as 200 million active users and
abrupt shutdown less than a year later~(\cite{vineDecline})).

@@

To that end, we can liken some aspects of online performance art on YouTube and Twitch
to street performance and busking,
allowing us to relate ostensibly new phenomena to robust scholarship.
As is the case with street performance,
online performance isn't restricted to classically trained professionals;
instead, performers (both online and in the streets) practice more experimentally,
drawing on candid interactions with their audiences,
which can change from one performance to the next.

This connection has purchase, but it also has gaps,
mostly owing to the unique nature of the internet~(\cite{miller2011understanding}).
For one thing, most video sharing websites keep videos
in perpetuity
--- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
The public that congregates around a performer, too, has a similar permanence that offline busking doesn't;
YouTube comments linger for years after they're made, forming in some sense ``networked publics''~(\cite{boyd2007youth}).
We must also account for the different nature of algorithmic enforcement of rules and policies, and
the near--perfect ability of these platforms to redirect revenue and
even an individual audience member's focus in an instant.

@@

\section*{Research Plan}
We propose to carry out qualitative research with major online content producers
on video \textit{streaming} sites such as Twitch and Periscope, and video \textit{sharing} sites like YouTube.
Much of this work will consist of ethnographic fieldwork, semi--structured interviews, and
open--ended exploration of the field site as we find parallels and divergences in the domain
of online performance art from street and other emergent formats of performance and entertainment.
A timeline of these phases is outlined here:

@@

We will recruit between 10 and 30 YouTube and Twitch content producers whose work is primarily content creation on their respective sites
--- in other words, people who have gone ``full--time''.
We will attempt to recruit approximately two--thirds of our participants from YouTube and one--third from Twitch, as
the former is significantly larger (and the community seemingly more diverse), suggesting
a need to get more varied perspectives from participants.

@@
